Learning and Reputation under Monopoly and Duopoly: \\A Competing Bandits Approach

### 2018-08-26 paper/competing_bandits.tex
uncommented a package to make it compile.
updated paper and appendix

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{amssymb}
\usepackage{natbib}
%% The amsthm package provides extended theorem environments
\usepackage{amsthm}
\usepackage{float}
\usepackage{sgame, tikz} % Game theory packages
\usepackage{caption} 
\usepackage{algorithm,algpseudocode}
\usepackage{makecell}
 \usepackage{multirow}
 \usepackage{graphicx}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\captionsetup{font=footnotesize}
\bibliographystyle{aaai}

@@

Unlike in the classic models in economics regarding the interplay between competition and innovation described in \citet{barro2004economic}, here we have no explicit R \& D cost. Rather, firms are not incentivized to experiment under duopoly because experimentation has an implicit relative \textit{reputational} cost compared to the myopic alternative and, in our model, consumers only care about the reputation of a firm. Thus, competition disincentivizes experimentation by firms precisely because competition forces firms to take actions in order to incentivize consumers to select them over their competitors and, as a result, by allowing a firm to be a temporary monopolist the firm can experiment without worrying about the reputational consequences. Our findings give an alternative intuition to the empirically documented inverse-U relationship between competition and innovation discussed in \citet{aghion2005competition}.

@@

\textbf{Overview} There are two firms and $T+2k$ agents where $k$ is the warm start, or the agents that each firm gets for free at the beginning of the game and $T$ is the number of rounds in the game. The timing of events is as follows:
\begin{enumerate}
\item At $t = 0$, the firms simultaneously commit to following a learning algorithm from a set of algorithms $\mathcal{A}$
\item Still at $t = 0$ we suppose that each firm gets $k$ agents as a "warm" start. The algorithm that the firm commits to makes $k$ rounds of progress and uses this information to initialize its information set. Additionally, the reputation score of the agents is initialized using the rewards from these $k$ rounds.
\item A new agent arrives each round (and lives for only one round), starting at $t = 1$, and chooses among 2 firms, given a reputation score for both firms.
\item The firm that is chosen selects an action $a_{t} \in A$ from a set of actions that is fixed across firms and rounds.
\item Both the agent and the firm observe the reward $r_t \in [0, 1]$ from the action. The agent reports this reward to the firm and the future agents and the reputation score for the chosen firm is updated.
\item Repeat 2-4 for $T$ rounds.
\end{enumerate}

Generally, the rewards are independent and identically distributed with a common prior.  For computational tractability we restrict our focus to Bernoulli-distributed rewards with Beta priors. Each firm faces a multi-armed bandit problem with no initial information \footnote{For algorithms that require some sort of prior to operate, such as Thompson Sampling, we use a "fake" prior of $Beta(\alpha=1,\beta=1)$}. We assume that the firms commit to a multi-armed bandit learning algorithm at the start of the world and that there are no informational spillovers from their competitors so that they can only learn from the agents that select them.

\noindent \textbf{Agents} We suppose that agents are homogenous, myopic, and non-strategic. The utility function for the agents is simply to maximize their reward in the one period in which they are alive. We suppose that agents do not attempt to manipulate the strategy of the firm nor do they take the strategies of the firm into account when choosing between the firms. In our model, each agent uses the average reward of past agents as a proxy for their expected utility. For simplicity, the reputation score, $R_{jt}$ is defined as a sliding window average\footnote{Another natural formulation would be to have exponential discounting of the past rewards. We believe that the results should be qualitatively similar to those presented here as in both formulations the more recent past matters more than the distant past.}.
\begin{center}
$R_{jt} = \frac{1}{M} \sum\limits_{i=1}^{M} r_{t_j-i}$
\end{center}

@@

Figure \ref{prelim_means} shows that the mean reputation ordering is as we would expect for the Needle In Haystack prior. The mean reputation ordering that we expect also holds for the other two priors and the results for those can be found in the GitHub appendix.

In the "competition game" the agents decide between the firms using the \textit{relative} reputation between the two firms. Namely, if firm $A$ has a higher reputation than firm $B$ agents will select firm $A$. A natural question to ask is whether looking at the mean performance is sufficient for understanding performance in the competition game. Figure \ref{rep_dist_nih} displays the reputation distribution at $t = 500$ for the same prior we reported the mean plot for previously. We see that the "naive" algorithms $DG$ and $DEG$ have a bi-modal reputation distribution whereas $TS$ does not. The intuition for this is that, for Needle In Haystack, $DG$ either finds the best arm or not. If it does, then since it engages in no purposeful exploration it will do better than any algorithm that engages in purposeful exploration over sufficiently many rounds. However, if it does not then it will get stuck on a bad arm and lose to $TS$ or $DEG$. In these cases its reputation may be substantially worse but for the competition game the difference between the reputation does not matter but only the binary comparison between them \footnote{This holds for our model and the decision rule of the agents, though the absolute difference may matter if, for instance, considering the SoftMax decision rule in \citet{mansour2018competing}}.

\begin{figure}
\caption{Reputation Distribution}
\includegraphics[scale=0.35]{figures/rep_distribution_nih}
\label{rep_dist_nih}
\caption*{\tiny{The plots contain a kernel density estimate of the reputation distribution at $t = 200$}}
\end{figure}

This motivates looking at the entire distribution of reputation difference between two algorithms. Figure \ref{ts_dg_rep_diff_nih} shows the distribution of reputation difference between $TS$ and $DG$ across $t$. This figure seems to confirm the intuition noted previously since the reputation distribution has its largest mass around the point just below 0 but it is skewed to the right even as $t$ gets large. Thus, the mean is not a representative statistic of the entire reputation difference distribution. As an alternative statistic for understanding the results of the competition game we introduce the \textit{relative reputation proportion} which looks at the proportion of simulations in which algorithm $A$ had at least as high of a reputation as algorithm $B$ for a fixed time $t$. This corresponds to running the bandit algorithms in isolation on the same instance and with the same realizations for $t$ rounds and then calculating the fraction of simulations at which an agent would select a firm playing $A$ over a firm playing $B$ at time $t$ \footnote{Though we do not discuss it here, one may be interested in if there is ever a case where, for sufficiently large $t$, we observe that $DEG >DG$ or $TS > DG$ according to the mean reputation but $DG > DEG$ or $DG > TS$ according to the relative reputation proportion. In the GitHub appendix we have results showing that, for Heavy Tail prior with $K=3$ we have that $DEG > DG$ according to the mean reputation but $DG > DEG$ according to the relative reputation proportion. Additionally, these results carry over to the competition game}.

@@

Figure \ref{relative_rep_plots} shows plots of the relative reputation proportion for $TS$ vs $DG$ on the Uniform and Needle In Haystack prior. For the Uniform prior we see that, in the early rounds, $DG > TS$ for the majority of the simulations but that, eventually, $TS > DG$. The intuition behind this is that, especially since the firms start with no substantive initial information, $TS$ does purposeful exploration in the early rounds in order to acquire information. However, this exploration leads to what we define as a \textit{relative reputation cost}. Namely, exploration leads to lower reputation relative to the myopic alternative. However, eventually the information acquired in the early rounds allows $TS$ to make better decisions and achieve a higher reputation, especially when the instance is "hard enough" so that $DG$ cannot trivially find the best arm. We define this as the \textit{relative reputation benefit} \footnote{We see similar behavior for the Heavy Tail prior}.

@@

Is exploration always costly? Figure \ref{relative_rep_plots} also shows that for the Needle In Haystack prior, $TS$ always does relatively better than $DG$. There are two contributing factors to this. First, $TS$ identifies the best arm faster in the Needle In Haystack prior than the Uniform prior so that there is a shorter time horizon where $TS$ needs to engage in purposeful exploration. Second, in the Needle In Haystack prior there are no "bad" arms as there may be in the Uniform prior since by construction all the arms except one in Needle In Haystack are the same. Thus, when $TS$ pulls a sub-optimal arm relative to its current information, the expected reward is the same as the greedy option that has not identified the best arm. However, with the Uniform prior, it is possible that the sub-optimal arm that is pulled has substantially lower expected reward relative to the greedy option.

@@

In this section we document our main results that can be summarized via the empirically documented inverted-U relationship between competition and innovation described in \citet{aghion2005competition}. We consider three separate settings for our model: permanent monopoly, temporary monopoly (incumbent vs entrant), and permanent duopoly. In the permanent monopoly and permanent duopoly cases we show that, under certain conditions, firms are incentivized to commit to $DG$ whereas in the temporary monopoly case firms are incentivized to commit to $TS$. We interpret the move from permanent monopoly to temporary monopoly and temporary monopoly to permanent duopoly as increases in competition. We believe this is reasonable and consistent with the empirical results which define increases in competition as increases in a measure of market power (1 - the Lerner Index). In our model we have no prices and no costs and so we interpret increases in "market power" as decreases in the number of rounds where a firm has to compete for agents.

@@

Since both firms enter at the same time the game is symmetric and we only need to compute the payoffs of three pairs of strategies. We do not report them in the figures, but when both firms play the same algorithm the expected market share is 50/50. Tables \ref{sim_nih}, \ref{sim_ht}, and \ref{sim_unif} report the results.
\begin{enumerate}
\item Under Needle In Haystack we have that $TS$ always does weakly better than $DEG$ and $DG$, regardless of the warm start
\item Under Heavy Tail we have that, for low warm start, $DEG, DG > TS$ but that for the higher warm start values we consider $TS > DEG, DG$. Interestingly, even under the relatively high warm start we see $DG > DEG$.
\item Under Uniform we have that $DEG, DG > TS$ for low warm start and $TS > DEG, DG$ for sufficiently high warm start.
\end{enumerate}

Looking at the relative reputation plots in Figure \ref{relative_rep_plots}, we can interpret fixing a warm start value as fixing the starting point on the relative reputation plots. The proportion of first rounds in the competition game that will go to a firm playing alg $A$ over a firm playing alg $B$ will correspond to the relative reputation proportion at time $k$. Combined with the fact that the EEOG values being relatively low across all warm starts \footnote{Interestingly, the EEOG also appears to be skewed to the right, similar to the distribution of reputation differences}, these observations seem to imply that understanding the performance of different algorithms in competition it is important to look at their performance for relatively small samples instead of asymptotically.

For the low values of the warm start we can see that $DG > TS$ for the Uniform prior but $TS > DG$ for the Needle In Haystack prior and this corresponds to what we see in the relative reputation proportion plots. The intuition for this result aligns with the intuition given in the discussion of the relative reputation plots. Namely, for low warm start values $TS$ incurs a relative reputation cost compared to $DG$ since it engages in purposeful exploration. Moving to a sufficiently large warm start so that the better algorithm has recovered from the relative reputation cost incurred via exploration leads to better algorithms winning. However, it appears that the warm starts required for this are unreasonably high, especially for the Uniform prior.

Focusing on reasonable warm starts, $DG > TS$ when it appears that better algorithms suffer an early relative reputational cost due to exploration and that both firms enter at the same time so that there is a "permanent duopoly."

\begin{table}[ht]
\centering
\caption{Duopoly Experiment Needle In Haystack} 
\begin{tabular}{rlll}
  \hline
 & k = 20 & k = 250 & k = 500 \\ 
  \hline
TS vs DG & \makecell{\textbf{0.64} $\pm$0.03\\ eeog \\ avg: 200\\ med: 27} & \makecell{\textbf{0.6} $\pm$0.03\\ eeog \\ avg: 370\\ med: 0} & \makecell{\textbf{0.64} $\pm$0.03\\ eeog \\ avg: 580\\ med: 121.5} \\ 
  TS vs DEG & \makecell{\textbf{0.57} $\pm$0.03\\ eeog \\ avg: 150\\ med: 14} & \makecell{\textbf{0.52} $\pm$0.03\\ eeog \\ avg: 460\\ med: 78.5} & \makecell{\textbf{0.56} $\pm$0.02\\ eeog \\ avg: 740\\ med: 627.5} \\ 
  DG vs DEG & \makecell{\textbf{0.46} $\pm$0.03\\ eeog \\ avg: 340\\ med: 128.5} & \makecell{\textbf{0.42} $\pm$0.02\\ eeog \\ avg: 650\\ med: 408} & \makecell{\textbf{0.42} $\pm$0.02\\ eeog \\ avg: 690\\ med: 466.5} \\ 
   \hline
\end{tabular}
\label{sim_nih}
\caption*{\tiny{The first line in each cell contains the average market share received by the firm playing Alg 1 (and the market share of Alg 2 is 1 - Alg 1 Market Share) as well as a 95 \% confidence band. For example, the cell in the top left indicates that TS gets on average 64\% of the market when played against DG. The next line contain the average and median effective end of game for this set of simulations.}}
\end{table}
% latex table generated in R 3.4.0 by xtable 1.8-2 package
% Thu Aug 16 13:13:00 2018
\begin{table}[ht]
\centering
\caption{Duopoly Experiment Heavy Tail} 
\begin{tabular}{rlll}
  \hline
 & k = 20 & k = 250 & k = 500 \\ 
  \hline
TS vs DG & \makecell{\textbf{0.29} $\pm$0.03\\ eeog \\ avg: 55\\ med: 0} & \makecell{\textbf{0.72} $\pm$0.02\\ eeog \\ avg: 570\\ med: 0} & \makecell{\textbf{0.76} $\pm$0.02\\ eeog \\ avg: 620\\ med: 98.5} \\ 
  TS vs DEG & \makecell{\textbf{0.3} $\pm$0.03\\ eeog \\ avg: 37\\ med: 0} & \makecell{\textbf{0.88} $\pm$0.01\\ eeog \\ avg: 480\\ med: 0} & \makecell{\textbf{0.9} $\pm$0.01\\ eeog \\ avg: 570\\ med: 113.5} \\ 
  DG vs DEG & \makecell{\textbf{0.62} $\pm$0.03\\ eeog \\ avg: 410\\ med: 7} & \makecell{\textbf{0.6} $\pm$0.02\\ eeog \\ avg: 790\\ med: 762} & \makecell{\textbf{0.57} $\pm$0.03\\ eeog \\ avg: 730\\ med: 608} \\ 
   \hline
\end{tabular}
\label{sim_ht}
\end{table}
% latex table generated in R 3.4.0 by xtable 1.8-2 package
% Thu Aug 16 13:13:01 2018
\begin{table}[ht]
\centering
\caption{Duopoly Experiment Uniform} 
\begin{tabular}{rlll}
  \hline
 & k = 20 & k = 250 & k = 500 \\ 
  \hline
TS vs DG & \makecell{\textbf{0.46} $\pm$0.03\\ eeog \\ avg: 230\\ med: 0} & \makecell{\textbf{0.52} $\pm$0.02\\ eeog \\ avg: 800\\ med: 754} & \makecell{\textbf{0.6} $\pm$0.02\\ eeog \\ avg: 910\\ med: 906.5} \\ 
  TS vs DEG & \makecell{\textbf{0.41} $\pm$0.03\\ eeog \\ avg: 180\\ med: 0} & \makecell{\textbf{0.51} $\pm$0.02\\ eeog \\ avg: 810\\ med: 734} & \makecell{\textbf{0.55} $\pm$0.02\\ eeog \\ avg: 970\\ med: 987} \\ 
  DG vs DEG & \makecell{\textbf{0.51} $\pm$0.03\\ eeog \\ avg: 470\\ med: 57.5} & \makecell{\textbf{0.48} $\pm$0.02\\ eeog \\ avg: 1000\\ med: 1088} & \makecell{\textbf{0.45} $\pm$0.02\\ eeog \\ avg: 1000\\ med: 1142} \\ 
   \hline
\end{tabular}
\label{sim_unif}
\end{table}

\textbf{Temporary Monopoly} We now consider asymmetries in the timing of entry so that one firm enters the market before the other and serves as a monopolist in the periods until the other firm enters. In terms of our model, this corresponds to setting $X = 200$. For this section we report results fixing $k = 20$. Does this lead the incumbent to commit to $TS$? How about the entrant?

Table \ref{ht_incum} shows the results of the simulations for this parameterization. We observe that, for the incumbent, $TS$ is the dominant strategy across all priors. In the appendix we include additional simulations where $X = 50$ and observe that $TS$ is not always a dominant strategy in this case. This result is robust across all of the priors that we considered so that for sufficiently large $X$, $TS$ is a dominant strategy for the incumbent where $X$ depends on the set of instances that we are considering.

>>>>

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{amssymb}
\usepackage{natbib}
%% The amsthm package provides extended theorem environments
\usepackage{amsthm}
\usepackage{float}
\usepackage{sgame, tikz} % Game theory packages
\usepackage{caption} 
\usepackage{algorithm,algpseudocode}
\usepackage{makecell}
 \usepackage{multirow}
 \usepackage{graphicx}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{finding}{Finding}
\captionsetup{font=footnotesize}
\bibliographystyle{aaai}
\frenchspacing

@@

Unlike in the classic models in economics regarding the interplay between competition and innovation described in \citet{barro2004economic}, here we have no explicit R \& D cost. Rather, firms are not incentivized to experiment under duopoly because exploration has an implicit relative \textit{reputational} cost compared to the greedy alternative and, in our model, consumers only care about the reputation of a firm. Thus, competition disincentivizes exploration by firms precisely because competition forces firms to take actions in order to incentivize consumers to select them over their competitors and, as a result, by allowing a firm to be a temporary monopolist the firm can experiment without worrying about the reputational consequences. Our findings give an alternative intuition to the empirically documented inverse-U relationship between competition and innovation discussed in \citet{aghion2005competition}.

@@

\textbf{Overview} There are two firms and $T+2k$ agents where $k$ is the warm start, or the agents that each firm gets for free at the beginning of the game and $T$ is the number of rounds in the game. The timing of events is as follows:
\begin{enumerate}
\item At $t = 0$, the firms simultaneously commit to following a learning algorithm from a set of algorithms $\mathcal{A}$
\item Still at $t = 0$ we suppose that each firm gets $k$ agents as a ``warm" start. The algorithm that the firm commits to makes $k$ rounds of progress and uses this information to initialize its information set. Additionally, the reputation score used by the agents is initialized using the rewards from these $k$ rounds.
\item A new agent arrives each round (and lives for only one round), starting at $t = 1$, and chooses the firm with the higher reputation at the time.
\item The firm that is chosen selects an action $a_{t} \in A$ from a set of actions that is fixed across firms and rounds.
\item Both the agent and the firm observe the reward $r_t \in [0, 1]$ from the action. The agent reports this reward to the firm and the future agents and the reputation score for the chosen firm is updated.
\item Repeat 2-4 for $T$ rounds.
\end{enumerate}

Generally, the rewards are independent and identically distributed with a common prior.  For computational tractability we restrict our focus to Bernoulli-distributed rewards with Beta priors. Each firm faces a multi-armed bandit problem with no initial information \footnote{For algorithms that require a prior to operate, such as Thompson Sampling, we use a "fake" prior of $Beta(\alpha=1,\beta=1)$}. We assume that the firms commit to a multi-armed bandit learning algorithm at the start of the world and that there are no informational spillovers from their competitors so that they can only learn from the agents that select them.

\noindent \textbf{Agents} We suppose that agents are homogenous, myopic, and non-strategic. The utility function for the agents is simply to maximize their reward in the one period in which they are alive. We suppose that agents do not attempt to manipulate the strategy of the firm, do not take the strategies of the firm into account when choosing between the firms, and do not condition their behavior on calendar time. In our model, each agent uses the average reward of past agents as a proxy for their expected utility. For simplicity, the reputation score, $R_{jt}$ is defined as a sliding window average\footnote{Another natural formulation would be to have exponential discounting of the past rewards. We believe that the results should be qualitatively similar to those presented here as in both formulations the more recent past matters more than the distant past.}.
\begin{center}
$R_{jt} = \frac{1}{M} \sum\limits_{i=1}^{M} r_{t_j-i}$
\end{center}

@@

Figure \ref{prelim_means} shows that the mean reputation ordering is as we would expect for the Needle In Haystack prior \footnote{The mean reputation ordering that we expect also holds for the other two priors and the results for those can be found in the supplementary material.}.

\begin{finding}
\textit{The mean performance of an algorithm is not a sufficient statistic for understanding the performance of an algorithm in competition. Looking at the distribution of reputation difference between algorithms we find that it tend is skewed to the right. Additionally we show that the mean performance of the algorithms is not always predictive of the competition game, even when both algorithms are given a large number of agents for free.}
\end{finding}

Is looking at the mean performance of between two algorithms sufficient for understanding what happens in competition? To answer this, we first look at the entire reputation distribution at $t = 500$ in Figure \ref{rep_dist_nih}. We see that the ``naive" algorithms $DG$ and $DEG$ have a bi-modal reputation distribution whereas $TS$ does not. The intuition for this is that, for Needle In Haystack, $DG$ either finds the best arm or not. If it does, then since it engages in no purposeful exploration it will do better than any algorithm that engages in purposeful exploration over sufficiently many rounds. However, if it does not then it will get stuck on a bad arm and lose to $TS$ or $DEG$. In these cases its reputation may be substantially worse but for the competition game the relative comparison between them is all that matters \footnote{This holds for our model and the decision rule of the agents, though the absolute difference may matter if, for instance, we consider the SoftMax decision rule in \citet{mansour2018competing}}.

To better capture the comparison between two algorithms we explicitly plot the distribution of the reputation difference between $TS$ and $DG$. This figure seems to confirm the intuition noted previously since the reputation distribution has its largest mass around the point just below 0 but it is skewed to the right even as $t$ gets large. Given the skewedness of the distribution, the mean is not a representative statistic of the entire reputation difference distribution. There are alternative statistics that we could consider such as the median or numerically calculating $\Pr(reputation(TS) - reputation(DG)) > 0$ using the estimated density, but instead we define a simple and natural statistic to help us reason about what happens under competition, the relative reputation proportion.

\begin{figure}
\caption{Reputation Distribution}
\includegraphics[scale=0.35]{figures/rep_distribution_nih}
\label{rep_dist_nih}
\caption*{\tiny{The plots contain a kernel density estimate of the reputation distribution at $t = 500$}}
\end{figure}

We define the \textit{relative reputation proportion} in order to better understand what happens under competition.

\begin{definition}
\textit{Relative Reputation Proportion} - the proportion of simulations in which algorithm $A$ had at least as high of a reputation as algorithm $B$ for a fixed time $t$
\end{definition}

\begin{finding}
\textit{Algorithms that engage in purposeful exploration can lead to implicit relative reputation costs compared to the greedy alternative in the earlier rounds. We observe this under the Uniform and Heavy Tail priors. However, it is not observed under the Needle In Haystack prior}
\end{finding}

The relative reputation statistic corresponds to running the bandit algorithms in isolation on the same instance and with the same realizations for $t$ rounds and then calculating the fraction of simulations at which an agent would select a firm playing $A$ over a firm playing $B$ at time $t$ \footnote{As further motivation for this statistic, one may be interested in if there is ever a case where, for sufficiently large $t$, we observe that $DEG >DG$ or $TS > DG$ according to the mean reputation but $DG > DEG$ or $DG > TS$ according to the relative reputation proportion. In the supplementary material we have results showing that, for Heavy Tail prior with $K=3$ we have that $DEG > DG$ according to the mean reputation but $DG > DEG$ according to the relative reputation proportion. Additionally, these results carry over to the competition game}.

@@

Figure \ref{relative_rep_plots} shows plots of the relative reputation proportion for $TS$ vs $DG$ on the Uniform and Needle In Haystack prior. For the Uniform prior we see that, in the early rounds, $DG > TS$ for the majority of the simulations but that, eventually, $TS > DG$. The intuition behind this is that, especially since the firms start with no substantive initial information, $TS$ does purposeful exploration in the early rounds in order to acquire information. However, eventually the information acquired from purposeful exploration in the early rounds allows $TS$ to make better decisions and achieve a higher reputation, especially when the instance is ``hard enough" so that $DG$ cannot trivially find the best arm. The early exploration leads to what we define as a \textit{relative reputation cost} and the eventual gain in reputation leads to what we define as a \textit{relative reputation benefit}.

\begin{definition}
\textit{Relative reputation cost and benefit} - the relative reputation loss from an algorithm that engages in purposeful exploration compared to the greedy alternative. It is an empirical definition based on the relative reputation proportion where the ``cost" regime is when the proportion $< 0.5$ and the ``benefit" regime is when the proportion is $> 0.5$ for the better algorithm. We call the priors where there is an early costly period followed by a later benefit period, relative reputation costly priors.
\end{definition}

@@

Is exploration always costly? Figure \ref{relative_rep_plots} also shows that for the Needle In Haystack prior, $TS$ always does relatively better than $DG$. There are two contributing factors to this. First, $TS$ identifies the best arm faster in the Needle In Haystack prior than the Uniform prior so that there is a shorter time horizon where $TS$ needs to engage in purposeful exploration. Second, in the Needle In Haystack prior there are no ``bad" arms as there may be in the Uniform prior since by construction all the arms except one in Needle In Haystack are the same. Thus, when $TS$ pulls a sub-optimal arm relative to its current information, the expected reward is the same as the greedy option that has not identified the best arm. However, with the Uniform prior, it is possible that the sub-optimal arm that is pulled has substantially lower expected reward relative to the greedy option. Thus, only the Uniform and Heavy Tail priors are relative reputation costly priors.

@@

In this section we document our main results that can be summarized via the empirically documented inverted-U relationship between competition and innovation described in \citet{aghion2005competition}. We consider three separate settings for our model: permanent monopoly, temporary monopoly (incumbent vs entrant), and permanent duopoly.

\begin{finding}
\textit{Under the priors that are relative reputation costly and with low warm start, the optimal strategies in the competition game are:
\begin{center}
\textbf{Permanent Monopoly} - $DG$ \\
\textbf{Temporary Monopoly} - $TS$ \\
\textbf{Permanent Duopoly} - $DG$ \\
\end{center}
The conditions on $TS$ being dominant under the temporary monopoly are that the incumbent is a temporary monopoly for sufficiently many periods.
}
\end{finding}

We interpret the move from permanent monopoly to temporary monopoly and temporary monopoly to permanent duopoly as increases in competition. We believe this is reasonable and consistent with the empirical results which define increases in competition as increases in a measure of market power (1 - the Lerner Index). In our model we have no prices and no costs and so we interpret increases in "market power" as decreases in the number of rounds where a firm has to compete for agents.

@@

Since both firms enter at the same time the game is symmetric and we only need to compute the payoffs of three pairs of strategies. We do not report them in the figures, but when both firms play the same algorithm the expected market share is 50/50.

\begin{finding}
\textit{For low warm starts, when exploration is relative reputation costly $DG >TS$ in competition, whereas when exploration is not relative reputation costly $TS > DG$. For sufficiently large warm starts, $TS$ is dominant across all priors.}
\end{finding}

Table \ref{sim_unif} shows the results under the Uniform prior. For the low values of the warm start we can see that $DG > TS$, but that for larger warm starts, $TS > DG$ \footnote{The same results hold for the Heavy Tail prior and are reported in the supplemental material}. Looking at the relative reputation plots in Figure \ref{relative_rep_plots}, we can interpret fixing a warm start value as fixing the starting point on the relative reputation plots. The proportion of first rounds in the competition game that will go to a firm playing alg $A$ over a firm playing alg $B$ will correspond to the relative reputation proportion at time $k$. The intuition for the result in competition aligns with the fact that for Uniform and Heavy Tail, $TS$ is relative reputation costly compared to $DG$ and so, for low warm starts, $TS$ has not yet recovered from the relative reputation loss due to purposeful exploration and so $DG$ beats $TS$. However, if we move the warm start sufficiently high so that $TS$ moves to the relative reputation benefit regime, $TS > DG$.

However, we do not always see that $DG > TS$ under some warm starts. Table \ref{sim_nih} shows the results for the Needle In Haystack prior where we see that $TS > DG$. The intuition for this is precisely that, under the Needle In Haystack, $TS$ is not relative reputation costly compared to $DG$ and thus we see that $TS > DG$.

Combining these results with the fact that the EEOG values are relatively low across all warm starts \footnote{Interestingly, the EEOG also appears to be skewed to the right, similar to the distribution of reputation differences}, these observations seem to imply that understanding the performance of different algorithms in competition it is important to look at their performance for relatively small samples instead of asymptotically.

% latex table generated in R 3.4.0 by xtable 1.8-2 package
% Thu Aug 16 13:13:01 2018
\begin{table}[ht]
\centering
\caption{Duopoly Experiment Uniform} 
\begin{tabular}{rlll}
  \hline
 & k = 20 & k = 250 & k = 500 \\ 
  \hline
TS vs DG & \makecell{\textbf{0.46} $\pm$0.03\\ eeog \\ avg: 230\\ med: 0} & \makecell{\textbf{0.52} $\pm$0.02\\ eeog \\ avg: 800\\ med: 754} & \makecell{\textbf{0.6} $\pm$0.02\\ eeog \\ avg: 910\\ med: 906.5} \\ 
  TS vs DEG & \makecell{\textbf{0.41} $\pm$0.03\\ eeog \\ avg: 180\\ med: 0} & \makecell{\textbf{0.51} $\pm$0.02\\ eeog \\ avg: 810\\ med: 734} & \makecell{\textbf{0.55} $\pm$0.02\\ eeog \\ avg: 970\\ med: 987} \\ 
  DG vs DEG & \makecell{\textbf{0.51} $\pm$0.03\\ eeog \\ avg: 470\\ med: 57.5} & \makecell{\textbf{0.48} $\pm$0.02\\ eeog \\ avg: 1000\\ med: 1088} & \makecell{\textbf{0.45} $\pm$0.02\\ eeog \\ avg: 1000\\ med: 1142} \\ 
   \hline
\end{tabular}
\label{sim_unif}
\end{table}

\begin{table}[ht]
\centering
\caption{Duopoly Experiment Needle In Haystack} 
\begin{tabular}{rlll}
  \hline
 & k = 20 & k = 250 & k = 500 \\ 
  \hline
TS vs DG & \makecell{\textbf{0.64} $\pm$0.03\\ eeog \\ avg: 200\\ med: 27} & \makecell{\textbf{0.6} $\pm$0.03\\ eeog \\ avg: 370\\ med: 0} & \makecell{\textbf{0.64} $\pm$0.03\\ eeog \\ avg: 580\\ med: 121.5} \\ 
  TS vs DEG & \makecell{\textbf{0.57} $\pm$0.03\\ eeog \\ avg: 150\\ med: 14} & \makecell{\textbf{0.52} $\pm$0.03\\ eeog \\ avg: 460\\ med: 78.5} & \makecell{\textbf{0.56} $\pm$0.02\\ eeog \\ avg: 740\\ med: 627.5} \\ 
  DG vs DEG & \makecell{\textbf{0.46} $\pm$0.03\\ eeog \\ avg: 340\\ med: 128.5} & \makecell{\textbf{0.42} $\pm$0.02\\ eeog \\ avg: 650\\ med: 408} & \makecell{\textbf{0.42} $\pm$0.02\\ eeog \\ avg: 690\\ med: 466.5} \\ 
   \hline
\end{tabular}
\label{sim_nih}
\caption*{\tiny{The first line in each cell contains the average market share received by the firm playing Alg 1 (and the market share of Alg 2 is 1 - Alg 1 Market Share) as well as a 95 \% confidence band. For example, the cell in the top left indicates that TS gets on average 64\% of the market when played against DG. The next line contain the average and median effective end of game for this set of simulations.}}
\end{table}
% latex table generated in R 3.4.0 by xtable 1.8-2 package
% Thu Aug 16 13:13:00 2018

\textbf{Temporary Monopoly} We now consider asymmetries in the timing of entry so that one firm enters the market before the other and serves as a monopolist in the periods until the other firm enters. In terms of our model, this corresponds to varying the value of $X$. For this section we report results fixing $k = 20$ and focus on the priors where we observed that, for this warm start, $DG > TS$ \footnote{However, we report the results for the Needle In Haystack prior in the supplementary material. The results are unsurprising, where $TS$ is the dominant strategy for both the entrant and incumbent.}. It seems clear that the temporary monopoly will get a larger market share regardless of what algorithm it plays, but is the incumbent incentivized to commit to $TS$?

\begin{finding}
\textit{Counterintuitively, allowing one firm to be a temporary monopolist for a sufficiently long time will induce that firm to commit to the better learning algorithm, $TS$, even if, under the permanent duopoly, they would commit to $DG$}
\end{finding}

In table \ref{ht_incum} we can see that, for $X = 200$, $TS$ is a dominant strategy for the incumbent and that $DG$ is the dominant strategy for the entrant on the Heavy Tail prior \footnote{The result for the Uniform is the same, though it takes a higher $X$ to induce the incumbent to play $TS$. However, while $DG > TS$ for the entrant, $DG$ does roughly the same as $DEG$ on the Uniform prior. The results are in the supplementary material.}. In the supplementary material we report the same experiment for different values of $X$ and show that our main result that, for sufficiently large $X$, $TS$ is the dominant strategy for the incumbent is robust across priors.

% latex table generated in R 3.4.0 by xtable 1.8-2 package
% Wed Aug 15 19:03:20 2018
\begin{table}[ht]
\centering
\caption{Temporary Monopoly Heavy Tail X = 200} 
\begin{tabular}{rlll}
  \hline
 & TS & DEG &  DG \\ 
  \hline
TS & \makecell{\textbf{0.003} $\pm$0.003\\Var:0.002\\ES:100\%} & \makecell{\textbf{0.083} $\pm$0.02\\Var:0.07\\ES:97\%} & \makecell{\textbf{0.17} $\pm$0.02\\Var:0.1\\ES:95\%} \\ 
  DEG & \makecell{\textbf{0.045} $\pm$0.01\\Var:0.03\\ES:92\%} & \makecell{\textbf{0.25} $\pm$0.02\\Var:0.1\\ES:75\%} & \makecell{\textbf{0.23} $\pm$0.02\\Var:0.1\\ES:78\%} \\ 
   DG & \makecell{\textbf{0.12} $\pm$0.02\\Var:0.08\\ES:88\%} & \makecell{\textbf{0.36} $\pm$0.03\\Var:0.2\\ES:76\%} & \makecell{\textbf{0.3} $\pm$0.02\\Var:0.1\\ES:64\%} \\ 
   \hline
\end{tabular}
\label{ht_incum}
\caption*{\tiny{The columns represent the strategy of the incumbent and the rows represent the strategy of the entrant. The first line in each cell contains the average market share for the entrant over $N=1000$ simulations as well as a 95\% confidence interval. The second line contains the sample variance of the observed market shares and the third line contains the fraction of simulations that ended up with one firm getting $> 90\%$ of the market. Note that smaller values in the table are better for the incumbent. Market shares are calculated as the fraction of users selecting a particular firm \textit{after} the entrant has already entered (i.e. the free rounds to firm 2 do not count towards the share)}}
\end{table}
